---
title: "SES Dashboard v1.0.2"
output: 
  flexdashboard::flex_dashboard:
    orientation: rows
    vertical_layout: fill
    theme: default
    social: [ "twitter", "facebook", "menu"]
    #source_code: embed
    logo: icon/logoCESIMAR.png
runtime: shiny
---

```{r setup, include=FALSE}
library(flexdashboard)
library(knitr)
library(stringr)
library(DT)
library(rpivotTable)
library(ggplot2)
library(plotly)
library(plyr)
library(dplyr)
library(highcharter)
library(ggvis)
library(leaflet)
library(dbscan)
library(factoextra)
library(shiny)
library(mregions)
library(sf)
library(zoo)
library(shinyWidgets)
library(tidyr)
library(reticulate)
library(WikidataQueryServiceR)
```


```{r}
data <- read.csv("ALLplatform.csv")
dataInstr <- read.csv("Instruments.csv")
datacensus <- read.csv("censusByCategories.csv")
datapapers <- read.csv("papers.csv")
dataBuoy <- read.csv("data_site_buoy.csv")
map_rendered <- reactiveVal(FALSE)
```

Dive Statistics
=====================================

Row
-------------------------------------

### Platform Analysis

```{r}
valueBox(paste("SES"),
         color = "warning",
         icon ="ion-stats-bars")
```

### Number of Platforms

```{r}
p1 <- data %>%
         group_by(claveU) %>%
         dplyr::summarise(count = n())

valueBox(length(p1$claveU),
         icon = "ion-ios-speedometer-outline" )

```

### Number of Dives

```{r}
valueBox(length(data$dive),
         icon = "ion-shuffle" )
```

### Max. Depth Recorded (m)

```{r}
valueBox(max(data$maxDepth),
         icon = "ion-arrow-graph-up-right" )
```

### Min. Depth Recorded (m)

```{r}
valueBox(min(data$minDepth),
         icon = "ion-arrow-graph-down-right" )
```

### Avg. Depth Recorded (m)

```{r}
valueBox(round(mean(data$avgDepth), digits = 2),
         icon = "ion-arrow-swap")
```

Row 
-------------------------------

### Sensors and Platforms

```{r}
renderDataTable({
  
  data <- data.frame(paste0("<a href='",dataInstr$uri,"'>",dataInstr$ClaveU,"</a>"),dataInstr$placementDate,
                      dataInstr$recoveryDate,dataInstr$type,paste0("<a href='",dataInstr$sensorURI,"'>",dataInstr$ID,"</a>"),dataInstr$maker,dataInstr$model)
  
  colnames(data) = c('Platform','Placement Data','Recovery Date','Type','Sensor ID','Manufactured by','Model')
  
  datatable(data, 
            options = list(scrollX = TRUE,pageLength = 10),
            escape = FALSE)
  
#  dataInstr
#datatable(dataInstr,
 #         rownames = T,
#          filter = "top",
#          options = list(pageLength = 10),list(scrollX = TRUE))%>% 
#          DT::formatStyle(columns = colnames(dataInstr), fontSize = '80%')

})
```

### Number of Dives By platform

```{r}
p2 <- data %>%
         group_by(claveU) %>%
         dplyr::summarise(count = n())

renderPlotly({
         plot_ly(x = ~p2$claveU,
                 y = ~p2$count,
                 color = "blue",
                 type = 'bar') %>%
        layout(xaxis = list(title = "Platform"),
        yaxis = list(title = 'Dives'))
})
```

Dive Analysis
========================================

Inputs {.sidebar}
-----------------------------------------------------------------------

```{r}
selectInput(inputId = "PlatformID", label = h3("Select Platform"),
                  choices = unique(data$claveU),
                  selected = "AMVI")

num <- reactive(input$PlatformID)

```

Row
-------------------------------
### Dive analysis (dives, dephts and temperatures)

```{r}

renderPlotly({
  
fields <- data  %>% select(dive,maxDepth,avgDepth,minDepth,duration,bottomTemp,surfaceTemp)  %>%
 filter(str_detect(data$claveU, num()))

vars <- setdiff(names(fields), "dive")

plots <- lapply(vars, function(var) {
  plot_ly(fields, x = ~dive, y = as.formula(paste0("~", var))) %>%
    add_lines(name = var) 
})
subplot(plots, nrows = length(plots), shareX = TRUE, titleX = TRUE) 
  
})
```

Buoy Data
========================================

Inputs {.sidebar}
-----------------------------------------------------------------------

```{r}
first_date <- min(as.Date(dataBuoy$timestamp))
last_date <- max(as.Date(dataBuoy$timestamp))


dateRangeInput(inputId = "dateRange", label = h3("Date Range"), 
               start = first_date, end = last_date,
               min = first_date, max = last_date,
               format = "yyyy-mm-dd")

HTML("<h3>Buoy Location</h3>")


output$mymap = renderLeaflet({
    leaflet() %>%
      addTiles() %>%
      setView(lng = -65.607, lat = -45.043, zoom = 12) %>%
      addMarkers(lng = -65.607, lat = -45.043)
})
leafletOutput('mymap')

```

Row {.tabset}
-------------------------------
### Isla Leones, Argentina

```{r}
htmltools::tags$p("*Wave height and Wind speed missing data values are approximated using a specialized function to enable visualization.")

renderPlotly({
  fields <- dataBuoy %>%
    select(timestamp, significant_wave_height_spotter, wind_speed_spotter, bottom_temperature_spotter, top_temperature_spotter)
  
  fields$timestamp <- as.POSIXct(fields$timestamp, format = "%Y-%m-%dT%H:%M:%OS", tz = "UTC")
  fields$formatted_date <- format(fields$timestamp, format = "%d/%m/%Y")

  filtered_fields <- fields %>% filter(timestamp >= input$dateRange[1] & timestamp <= input$dateRange[2])

  filtered_fields$bottom_temperature_spotter <- na.approx(filtered_fields$bottom_temperature_spotter)
  filtered_fields$top_temperature_spotter <- na.approx(filtered_fields$top_temperature_spotter)

  mean_temp <- (filtered_fields$bottom_temperature_spotter + filtered_fields$top_temperature_spotter) / 2
  
  filtered_fields$significant_wave_height_spotter <- approx(filtered_fields$timestamp, 
                                                          filtered_fields$significant_wave_height_spotter,
                                                          xout = filtered_fields$timestamp)$y
  
  filtered_fields$wind_speed_spotter <- approx(filtered_fields$timestamp, 
                                                          filtered_fields$wind_speed_spotter,
                                                          xout = filtered_fields$timestamp)$y

  plot <- plot_ly(filtered_fields, x = ~timestamp) %>%
    add_trace(y = ~top_temperature_spotter, type = "scatter", mode = "lines", name = "Top Temp") %>%
    add_trace(y = mean_temp, type = "scatter", mode = "lines", name = "Mean Temp", line = list(dash = 'solid')) %>%
    add_trace(y = ~bottom_temperature_spotter, type = "scatter", mode = "lines", name = "Bottom Temp")


  plot_temperatures <- plot_ly(filtered_fields, x = ~timestamp) %>%
    add_trace(y = ~top_temperature_spotter, type = "scatter", mode = "lines", name = "Top Temp") %>%
    add_trace(y = mean_temp, type = "scatter", mode = "lines", name = "Mean Temp", line = list(dash = 'solid')) %>%
    add_trace(y = ~bottom_temperature_spotter, type = "scatter", mode = "lines", name = "Bottom Temp")
  
  plot_temperatures<- plot_temperatures %>% layout(
    yaxis = list(title = "Temperatures (°C)", range = c(0, max(filtered_fields$significant_wave_height_spotter))))

  plot_wave_height <- plot_ly(filtered_fields, x = ~timestamp) %>%
    add_trace(y = ~significant_wave_height_spotter, type = "scatter", mode = "lines", name = "Wave Height")
  
  plot_wave_height<- plot_wave_height %>% layout(
    yaxis = list(title = "Wave height (mts)", range = c(0, max(filtered_fields$significant_wave_height_spotter))))

  plot_wind_speed <- plot_ly(filtered_fields, x = ~timestamp) %>%
    add_trace(y = ~wind_speed_spotter, type = "scatter", mode = "lines", name = "Wind Speed")
  
  plot_wind_speed<- plot_wind_speed %>% layout(
    yaxis = list(title = "Wind speed (km/h)", range = c(0, max(filtered_fields$significant_wave_height_spotter))))

  subplot(plot_temperatures, plot_wave_height, plot_wind_speed, nrows = 3, shareX = TRUE, titleX = TRUE, titleY = TRUE)%>%
    layout(heights = "50%")
})
```
### Otra pestaña


```{r}
source_python('queryListOfSpeciesWikidata.py')
species <- queryListOfSpecies()

selectInput(inputId = "Specie", label = h5("Species from BiGe-Onto Endpoint"),
                  choices = unique(species),
            selected = "Mirounga leonina")

selectionSpecie <- reactive({input$Specie})
source_python('querySpecieWikidata.py')

y <- reactive({
  querySpecieResult <- querySpecie(selectionSpecie())
  return(querySpecieResult)
})

# Continuar aquí. Imrimir el valor de y() para ver la salida. Trabajar en la tabla





renderDataTable({
  results <- y()
  
    if (nrow(results) == 0) {
    # Si el DataFrame está vacío, crear un DataFrame vacío con las mismas columnas
      results <- data.frame(
        'Wikidata item' = character(0),
        'scientific_name' = character(0),
        'common_name' = character(0),
        'status' = character(0),
        'rangemap' = character(0),
        'length' = character(0),
        'life_expectency' = character(0),
        'height' = character(0),
        'mass' = character(0)
      )
    }
  
  m<<- data.frame(paste0("<a href='",results$item,"'>",results$item,"</a>"),results$scientific_name,results$common_name,paste0("<a href='",results$status,"'>",results$status,"</a>"), paste0("<img src='",results$rangemap,"' height=42></img>"),results$length,results$life_expectency, results$height,results$mass)


colnames(m) = c('Wikidata item','scientific_name','common_name','status','rangemap','length','life_expectency','height','mass')

DT::datatable(m, class = 'cell-border stripe',fillContainer = T,options = list(autoWidth = TRUE,scrollX = TRUE, dom = 't'), escape = FALSE)
})

```
Platform Trips
========================================

Inputs {.sidebar}
-----------------------------------------------------------------------

```{r}
selectInput(inputId = "PlatformTrip", label = h3("Select Platform"),
                  choices = unique(data$claveU),
                  selected = "ANIF")

platform <- reactive(input$PlatformTrip)

```

Row
-------------------------------
### Trips made by all platforms

```{r}
 renderLeaflet({
  
  trips <- data  %>% select(lati,longi,maxDepth,duration,bottomTemp,surfaceTemp)  %>%
          filter(str_detect(data$claveU, platform()))
  
  map <- leaflet(trips) %>% addTiles(group = "OpenStreetMap")
  map = map %>% addCircleMarkers(~longi, ~lati, radius = 1, 
       popup = paste("<b>Max Depht</b> (meters): ", trips$MaxDepth,
                                 "<b>Dive Duration</b> (seconds): ",trips$duration, 
                                 "<b>Bottom temperature</b> (°C): ",trips$bottomTemp,
                                 "<b>Surface temperature</b> (°C): ",trips$surfaceTemp,
                                 sep = "<br/>"),
        weight = 2, opacity = 0.5, fill = TRUE, fillOpacity = 0.2, color = "red") %>%
  addPolylines(data=trips, lng = ~longi, lat = ~lati, stroke = TRUE, color = "green", weight = 2)%>% 
  addTiles() %>% 
  addProviderTiles(providers$Esri.NatGeoWorldMap)%>%
  addMeasure()%>%
  addMeasure(  position = "bottomleft",  primaryLengthUnit = "kilometers",  primaryAreaUnit = "sqkilometers",  activeColor = "#3D535D",  completedColor = "#7D4479",  localization = "en")
  
})
```


Cluster Analysis
========================================

Column {.sidebar}
-----------------------------------------------------------------------
Two important parameters are required for DBSCAN:

- epsilon (“eps”): The parameter eps defines the radius of neighborhood around a point x. 

- MinPts: Is the minimum number of neighbors within “eps” radius.

```{r}
selectInput("n_points", label = "MinPts:",
            choices = c(5, 10, 12, 15, 20), selected = 12)
points <- reactive(input$n_points)

sliderInput("eps_adjust", label = "eps:",
           min = 0.01, max = 0.20, value = 0.15, step = 0.02)
epsvalue <- reactive(input$eps_adjust)

```

row
-------------------------------------

### Cluster analysis using dbscan

```{r}
renderPlotly({
  
  locs <- dplyr::select(data,lati,longi)
  
  #scalling the data points.
  locs.scaled = scale(locs,center = T,scale = T)
  # esta sección es la que rompe los mapas
  db = dbscan::dbscan(locs.scaled,eps=epsvalue(),minPts = as.numeric(points()))

  hclust_result <- hclust(dist(locs.scaled))
  clusters <- cutree(hclust_result, k = as.numeric(points()))

  plot_data <- data.frame(lati = data$lati, longi = data$longi, Cluster = factor(clusters))

  fig <- plot_ly(data = plot_data, x = ~lati, y = ~longi, color = ~Cluster) %>%
      add_markers()
    
  fig
  
  factoextra::fviz_cluster(db,locs.scaled,stand = F,ellipse = T,geom = "point")
})
```

Census Statistics
========================================

column
-------------------------------------
    
### Census by year and SES category

```{r}
renderPlotly({
plot_ly(datacensus, x=~datacensus$year, y=~datacensus$total, color=~datacensus$category, type="bar") %>% layout(barmode = "stack", title = "Census by year and SES category",
         xaxis = list(title = "Years"),
         yaxis = list(title = "Amount"))
})
```


Bibliography
========================================

Row
-------------------------------
### Papers associated with each platform

```{r}

renderDataTable({
  
  data <- data.frame(paste0("<a href='",datapapers$platform,"'>",datapapers$claveU,"</a>"),datapapers$title,
                     paste0("<a href='",datapapers$DOI,"'>",datapapers$DOI,"</a>"))
  
  colnames(data) = c('Platform','Title','DOI')
  
  datatable(data, 
            options = list(scrollX = TRUE,pageLength = 25),
            escape = FALSE)


})
```

About
========================================
<font size="5">SES Dashboard: A linked data platform to manage data on Southern Elephant Seals</font> 

SES Dashboard is a linked data platform developed to manage information of Southern Elephant Seals (Mirounga leonina, Linnaeus 1758) from Península Valdés colony (Patagonia, Argentina), collected over more than two decades of research.

This platform provides support for decision-making related to the conservation of the species inhabiting the Southwestern Atlantic Ocean. Also, the version 2021 published following [Linked Open Data (LOD) principles](https://www.w3.org/wiki/LinkedData), which allows to increase the intended audience by reusing of data, so the range of possibilities is broadened.

<font size="5">Packages used</font> 

The application is built with the [Shiny](http://shiny.rstudio.com) framework for the [R programming language](https://www.r-project.org/). Access to endpoints is done through the [SPARQL](https://cran.r-project.org/web/packages/SPARQL/SPARQL.pdf) package. The application layout is produced with the [flexdashboard](http://rstudio.github.io/flexdashboard/index.html) package, and the maps use [Leaflet.js](http://leafletjs.com/), [Highcharts](http://www.highcharts.com/), and [ggplot2](http://ggplot2.org/), all accessed through their corresponding R packages.  

<font size="5">Availability for use and reuse</font> 

Code for the application is available [here](https://github.com/marcosdzarate/ses-dashboard). I welcome feedback and suggestions! for contact information or [connect with me on Twitter](https://twitter.com/MarcosdZarate84).  